---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      fig.width = 12,
                      fig.height = 6,
                      fig.path = "README_figures/README-")
```


# wtkapi

The goal of wtkapi is to get data from the [Gridded Atmospheric Wind Integration National Dataset Toolkit](https://www.nrel.gov/grid/wind-toolkit.html) in tidy format to leverage technical analysis of atmospheric data.

It is required to sign up for an API key on the [NREL Developer Network](https://developer.nrel.gov/). Please also consider the API call rate limits.


## Installation

You can install the development version of wtkapi from GitHub with:

``` r
# install.packages("devtools")
devtools::install_github("floresfdev/wtkapi")
```

At the moment this package is heavy on dependencies, so in order to have more control over your environment, avoid install/upgrade dependencies with:

``` r
devtools::install_github("floresfdev/wtkapi", dependencies = FALSE, upgrade_dependencies = FALSE)
```


## Example

This is a basic example which shows you how to get data:

```{r}
library(dplyr)
library(lubridate)
library(ggplot2)
library(glue)
library(wtkapi)

# API parameters
endpoint <- "https://developer.nrel.gov/api/hsds/"
host <- "/nrel/wtk-us.h5"
api_key <- readr::read_lines(here::here("api_key.txt"))

# Get the info of the available datasets
datasets <- get_datasets(endpoint, host, api_key)

# Info to select the year 2007, with 1-hour resolution data
datetime_info <-
    compute_datetime_info(datetime_from = "2007-01-01 00:00:00",
                          datetime_to = "2007-12-31 23:00:00",
                          datetime_step = 1)
                          
# Info to select the location: New York
latitude <- 40.7128
longitude <- -74.0059

# Info to select the datasets as columnar data
datasets$title

dataset_titles <- c("temperature_10m", "windspeed_10m", "winddirection_10m",
                        "temperature_40m", "windspeed_40m", "winddirection_40m",
                        "temperature_80m", "windspeed_80m", "winddirection_80m")

# Get data from the API
wind_temp_2007 <- 
    get_dataset(datasets,
                api_key,
                dataset_titles,
                datetime_info,
                latitude,
                longitude)

# Number of rows: 8760 = 365 days * 24 hours
nrow(wind_temp_2007)

# Check first and last records
head(wind_temp_2007)
tail(wind_temp_2007)

# Plots

plot_subtitle <- glue("2007, New York (lat {latitude}, lon {longitude})")

## Daily min/max temperature at 10mts
wind_temp_2007 %>%
    ## Convert temperature from Kelvin to Celsius
    mutate(temperature_10m = temperature_10m - 273.15) %>%
    ## Get the min and max temperature for each day
    mutate(date = as_date(datetime)) %>%
    group_by(date) %>%
    summarize(min_temperature_10m = min(temperature_10m),
              max_temperature_10m = max(temperature_10m)) %>%
    ## Plot timeseries
    ggplot() +
    geom_line(aes(x = date,
                  y = min_temperature_10m),
              color = "blue") +
    geom_line(aes(x = date,
                  y = max_temperature_10m),
              color = "red") +
    labs(x = "Date",
         y = "Temperature (°C)",
         title = "Daily min/max temperature at 10mts",
         subtitle = plot_subtitle,
         caption = "Blue = daily min  -  Red = daily max") +
    theme_classic()


## Daily average wind speed at 80mts
wind_temp_2007 %>%
    ## Get the average of windspeed for each day
    mutate(date = as_date(datetime)) %>%
    group_by(date) %>%
    summarize(avg_windspeed_80m = mean(windspeed_80m)) %>%
    ## Plot timeseries
    ggplot() +
    geom_line(aes(x = date,
                  y = avg_windspeed_80m)) +
    labs(x = "Date",
         y = "Wind speed (m/s)",
         title = "Daily average wind speed at 80mts",
         subtitle = plot_subtitle) +
    theme_classic()


## Histogram of wind speed at 80 mts
ggplot(wind_temp_2007) +
    geom_histogram(aes(x = windspeed_80m),
                   fill = "white",
                   color = "black") +
    labs(x = "Wind speed (m/s)",
         y = "Count",
         title = "Histogram of wind speed at 80 mts",
         subtitle = plot_subtitle) +
    theme_classic()

```

